The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-03-01 article/article.tex
improved conclusion

Consider the slightly more Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, we put the ball into the left-most bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers. Refer to Appendix \ref{sec:fibonacci} for the definition.}
\begin{align*}
\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.

@@

\section{Conclusion}
\label{sec:conclusion}
BiB games have many applications in Computer Science and have been studied extensively. The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball. The classical result for this scheme upper-bounds the maximum load with 
\begin{align*}
\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
\end{align*}
w.h.p.
In this paper we have presented two different miltiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
\begin{align*}
\load = \frac{m}{n} + \frac{\ln\ln n}{\ln d} + O(1)
\end{align*}
w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better maximum load of
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds in detail.

>>>>

Consider the slightly more Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, we put the ball into the left-most bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers. Refer to Appendix \ref{sec:fibonacci} for the definition.}
\begin{align*}
\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p.\ for all $m$. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.

@@

\section{Conclusion}
\label{sec:conclusion}
BiB games have many applications in Computer Science and have been studied extensively. The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball. The classical result for this scheme upper-bounds the gap with \cite{RS98}
\begin{align*}
\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
\end{align*}
w.h.p.
In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
\begin{align*}
\delta = \frac{\ln\ln n}{\ln d} + O(1)
\end{align*}
w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better gap of
\begin{align*}
\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
